---
title: "Project of CIML"
author: "Sergio Blanco Piñeiro"
header-includes:
date: "`r Sys.Date()`"
output:
  pdf_document: default
  code_folding: "hide" 
fontsize: 12pt
institute: "Universidad Carlos III de Madrid"
department: "Department of Economics"
short-author: "CIML (Spring 2023)"
short-date: "03/30/2023"
short-institute: "UC3M Economics"
subtitle: "The Sensitivity of an Empirical Model of Married Women's Hours of Work to Economic
  and Statistical Assumptions"
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
suppressWarnings(suppressMessages(library(tidyverse)))
library(haven)
library(tidyverse)
library(GGally)
library(extrafont)
library(dplyr)
library(marginaleffects)
library(glmnet)
library(MASS)
library(rpart)
library(rpart.plot)
library(randomForest)
library(class)
library(e1071)
library(stats)
library(Metrics) 
library(neuralnet)
library(kableExtra)
library(sandwich)
library(stargazer)
theme_set(theme_bw() + theme(text=element_text(family="Palatino", size=20)))
fig.width=6
fig.asp=0.618
fig.align="center"
out.width="75%"

```

## Data loading and visualization

```{r 1, include=FALSE, echo=TRUE}
setwd("C:\\Users\\sbp99\\Desktop\\Causal Inference and Machine Learning\\Project")
data_frame <- read_dta("C:\\Users\\sbp99\\Desktop\\Causal Inference and Machine Learning\\Project\\mroz.dta")
```

We will compute some basic descriptive statistics for the variables of this data set. In the following boxplots we can see that we have a Sample Selection issue since the Education, Age, Experience, Family income excluding women earnings and number of kids differ across women that participate and not on the Labour Market.

```{r 2, include=FALSE, echo=TRUE}
attach(data_frame)
head(data_frame)
str(data_frame)
summary(data_frame)
sapply(data_frame, sd)
by(data_frame, inlf, summary)

```

```{r 3, include=TRUE, echo=TRUE}
boxplot(age~inlf, ylab="Age", xlab="Participation in the Labor Market",col = "blue")
boxplot(educ~inlf, ylab="Education", xlab="Participation in the Labor Market",col = "blue")
boxplot(exper~inlf, ylab="Experience", xlab="Participation in the Labor Market",col = "blue")
boxplot(nwifeinc~inlf, ylab="Family income exluding women earnings", xlab="Participation in the Labor Market",col = "blue")
boxplot(kidsge6~inlf, ylab="Number of kids between 6 to 18 years old", xlab="Participation in the Labor Market",col = "blue")
boxplot(kidslt6~inlf, ylab="Number of kids < 6 years old", xlab="Participation in the Labor Market",col = "blue")
```

Lets compute a naive regression for the subpopulation of women that participate in the labor market. We will estimate the following equation: $log(Wage)_i=\beta_0 + \beta_1*Age_i + \beta_2*Education_i + \beta_3*Experience_i + \beta_4*{Experience_i}^2 + \beta_5*City_i+ \epsilon_i$ where we have as explanatory variables woman's age, education, labor market experience and its square, and the binary variable on living in a large city and the log of wages as a dependent variable. After we estimate this naive regression as our objective in this part is to generate an imputed log(Wages) (called, i_log(Wages)) Once we achieve that we will use that prediction in order to impute the i_log(Wage) for the non-participant women and us the prediction as the i_log(Wage) for the participant women. Even though we imputed this log(Wage) by applying the Heckman two-step-procedure we will not use this i_log(Wages) as this method accounts for Sample Selection.

## Imputation of log(Wage)

```{r 4, include=FALSE, echo=TRUE}
attach(data_frame)
participants_data_frame<-subset(data_frame, inlf == 1)
non_participants_data_frame<-subset(data_frame, inlf == 0)
ols_subsample<-lm(lwage~age+educ+exper+expersq+city, data =participants_data_frame )
summary(ols_subsample)
prediction_1<-predict(ols_subsample, newdata=non_participants_data_frame)
data_frame$lww1<-ifelse(data_frame$inlf==1,data_frame$lwage,prediction_1)
data_frame$kids<-(data_frame$kidslt6  + data_frame$kidsge6)
```

```{r 5, include=FALSE, echo=TRUE}
attach(data_frame)

#Interaction with city=1{if i lives in a big city}
data_frame$city_kids<-city*kids
data_frame$city_age<-city*age
data_frame$city_educ<-city*educ
data_frame$city_hushrs<-city*hushrs
data_frame$city_husage<-city*husage
data_frame$city_huseduc<-city*huseduc
data_frame$city_huswage<-city*huswage
data_frame$city_unem<-city*unem
data_frame$city_mtr<-city*mtr
data_frame$city_faminc<-city*faminc
data_frame$city_motheduc<-city*motheduc
data_frame$city_fatheduc<-city*fatheduc
data_frame$city_unem<-city*unem
data_frame$city_exper<-city*exper
data_frame$city_expersq<-city*expersq

#Square of covariates
data_frame$kids_sq<-kids*kids
data_frame$age_sq<-age*age
data_frame$educ_sq<-educ*educ
data_frame$hushrs_sq<-hushrs*hushrs
data_frame$husage_sq<-husage*husage
data_frame$huseduc_sq<-huseduc*huseduc
data_frame$huswage_sq<-huswage*huswage
data_frame$faminc_sq<-faminc*faminc
data_frame$mtr_sq<-mtr*mtr
data_frame$motheduc_sq<-motheduc*motheduc
data_frame$fatheduc_sq<-fatheduc*fatheduc
data_frame$unem_sq<-unem*unem

#Covariates to the power of 3 
data_frame$kids_3<-kids*kids*kids
data_frame$age_3<-age*age*age
data_frame$educ_3<-educ*educ*educ
data_frame$hushrs_3<-hushrs*hushrs*hushrs
data_frame$husage_3<-husage*husage*husage
data_frame$huseduc_3<-huseduc*huseduc*huseduc
data_frame$huswage_3<-huswage*huswage*huswage
data_frame$faminc_3<-faminc*faminc*faminc
data_frame$mtr_3<-mtr*mtr*mtr
data_frame$motheduc_3<-motheduc*motheduc*motheduc
data_frame$fatheduc_3<-fatheduc*fatheduc*fatheduc
data_frame$unem_3<-unem*unem*unem

#Covariates to the power of 4
data_frame$kids_4<-kids*kids*kids*kids
data_frame$age_4<-age*age*age*age
data_frame$educ_4<-educ*educ*educ*educ
data_frame$hushrs_4<-hushrs*hushrs*hushrs*hushrs
data_frame$husage_4<-husage*husage*husage*husage
data_frame$huseduc_4<-huseduc*huseduc*huseduc*huseduc
data_frame$huswage_4<-huswage*huswage*huswage*huswage
data_frame$faminc_4<-faminc*faminc*faminc*faminc
data_frame$mtr_4<-mtr*mtr*mtr*mtr
data_frame$motheduc_4<-motheduc*motheduc*motheduc*motheduc
data_frame$fatheduc_4<-fatheduc*fatheduc*fatheduc*fatheduc
data_frame$unem_4<-unem*unem*unem*unem

#Covariates to the power of 5
data_frame$kids_5<-kids*kids*kids*kids*kids
data_frame$age_5<-age*age*age*age*age
data_frame$educ_5<-educ*educ*educ*educ*educ
data_frame$hushrs_5<-hushrs*hushrs*hushrs*hushrs*hushrs
data_frame$husage_5<-husage*husage*husage*husage*husage
data_frame$huseduc_5<-huseduc*huseduc*huseduc*huseduc*huseduc
data_frame$huswage_5<-huswage*huswage*huswage*huswage*huswage
data_frame$faminc_5<-faminc*faminc*faminc*faminc*faminc
data_frame$mtr_5<-mtr*mtr*mtr*mtr*mtr
data_frame$motheduc_5<-motheduc*motheduc*motheduc*motheduc*motheduc
data_frame$fatheduc_5<-fatheduc*fatheduc*fatheduc*fatheduc*fatheduc
data_frame$unem_5<-unem*unem*unem*unem*unem

#Interaction with faminc
data_frame$faminc_kids<-faminc*kids
data_frame$faminc_age<-faminc*age
data_frame$faminc_educ<-faminc*educ
data_frame$faminc_hushrs<-faminc*hushrs
data_frame$faminc_husage<-faminc*husage
data_frame$faminc_huseduc<-faminc*huseduc
data_frame$faminc_huswage<-faminc*huswage
data_frame$faminc_mtr<-faminc*mtr
data_frame$faminc_motheduc<-faminc*motheduc
data_frame$faminc_fatheduc<-faminc*fatheduc
data_frame$faminc_unem<-faminc*unem
data_frame$faminc_exper<-faminc*exper
data_frame$faminc_expersq<-faminc*expersq

#Interaction with unem
data_frame$unem_kids<-unem*kids
data_frame$unem_age<-unem*age
data_frame$unem_educ<-unem*educ
data_frame$unem_hushrs<-unem*hushrs
data_frame$unem_husage<-unem*husage
data_frame$unem_huseduc<-unem*huseduc
data_frame$unem_huswage<-unem*huswage
data_frame$unem_mtr<-unem*mtr
data_frame$unem_motheduc<-unem*motheduc
data_frame$unem_fatheduc<-unem*fatheduc
data_frame$unem_exper<-unem*exper
data_frame$unem_expersq<-unem*expersq

#Interaction with huswage
data_frame$huswage_kids<-huswage*kids
data_frame$huswage_age<-huswage*age
data_frame$huswage_educ<-huswage*educ
data_frame$huswage_hushrs<-huswage*hushrs
data_frame$huswage_husage<-huswage*husage
data_frame$huswage_huseduc<-huswage*huseduc
data_frame$huswage_mtr<-huswage*mtr
data_frame$huswage_motheduc<-huswage*motheduc
data_frame$huswage_fatheduc<-huswage*fatheduc
data_frame$huswage_exper<-huswage*exper
data_frame$huswage_expersq<-huswage*expersq

#Interaction with fatheduc
data_frame$fatheduc_kids<-fatheduc*kids
data_frame$fatheduc_age<-fatheduc*age
data_frame$fatheduc_educ<-fatheduc*educ
data_frame$fatheduc_hushrs<-fatheduc*hushrs
data_frame$fatheduc_husage<-fatheduc*husage
data_frame$fatheduc_huseduc<-fatheduc*huseduc
data_frame$fatheduc_mtr<-fatheduc*mtr
data_frame$fatheduc_motheduc<-fatheduc*motheduc
data_frame$fatheduc_exper<-fatheduc*exper
data_frame$fatheduc_expersq<-fatheduc*expersq


#Interaction with motheduc
data_frame$motheduc_kids<-motheduc*kids
data_frame$motheduc_age<-motheduc*age
data_frame$motheduc_educ<-motheduc*educ
data_frame$motheduc_hushrs<-motheduc*hushrs
data_frame$motheduc_husage<-motheduc*husage
data_frame$motheduc_huseduc<-motheduc*huseduc
data_frame$motheduc_mtr<-motheduc*mtr
data_frame$motheduc_exper<-motheduc*exper
data_frame$motheduc_expersq<-motheduc*expersq

#Interaction with educ
data_frame$educ_kids<-educ*kids
data_frame$educ_age<-educ*age
data_frame$educ_hushrs<-educ*hushrs
data_frame$educ_husage<-educ*husage
data_frame$educ_huseduc<-educ*huseduc
data_frame$educ_mtr<-educ*mtr
data_frame$educ_exper<-educ*exper
data_frame$educ_expersq<-educ*expersq

#Interaction with age
data_frame$age_kids<-age*kids
data_frame$age_hushrs<-age*hushrs
data_frame$age_husage<-age*husage
data_frame$age_huseduc<-age*huseduc
data_frame$age_mtr<-age*mtr
data_frame$age_exper<-age*exper
data_frame$age_expersq<-age*expersq

#Interaction with hushrs
data_frame$hushrs_kids<-hushrs*kids
data_frame$hushrs_husage<-hushrs*husage
data_frame$hushrs_huseduc<-hushrs*huseduc
data_frame$hushrs_mtr<-hushrs*mtr
data_frame$hushrs_exper<-hushrs*exper
data_frame$hushrs_expersq<-hushrs*expersq

#Interaction with husage
data_frame$husage_kids<-husage*kids
data_frame$husage_huseduc<-husage*huseduc
data_frame$husage_mtr<-husage*mtr
data_frame$husage_exper<-husage*exper
data_frame$husage_expersq<-husage*expersq

#Interaction with huseduc
data_frame$huseduc_kids<-huseduc*kids
data_frame$huseduc_mtr<-huseduc*mtr
data_frame$huseduc_exper<-huseduc*exper
data_frame$huseduc_expersq<-huseduc*expersq
```

```{r 6, include=FALSE, echo=TRUE}

set.seed(1)
sample<-list()
train<-list()
test<-list()

for (i in seq(1,5)){

 sample[[i]]<-sample(c(TRUE, FALSE), nrow(data_frame), replace=TRUE, prob=c(0.8,0.2))
 train[[i]]  <- data_frame[sample[[i]], ]
 test[[i]]  <- data_frame[!sample[[i]], ]
}
  


```

## First Step estimation

Once, we have created our variable i_log(Wage). Lets estimate the following equation $D_i:= 1\{inlf_i=1\}=F(\textbf{Z}_i^{T}\gamma)$ using a set of covariates $\textbf{Z}_i$. In order to estimate this model we will use firstly a Linear Probability model, Logit model and a Probit model using a naive regression with a small set of covariates. But before we estimate the previous model, we will create a HD setting.

```{r 7, include=FALSE, echo=TRUE}

#### Naive Model Fitting

#Naive Logit

logit_<-list()
for (i in seq(1,length(train))){
logit_[[i]]<-glm(inlf~kids + age + educ + nwifeinc + unem + city, data = train[[i]], family = "binomial")
print(logit_[[i]])
}

#Naive Probit

probit_<-list()
for (i in seq(1,length(train))){
probit_[[i]]=glm(inlf~kids + age + educ + nwifeinc + unem + city, data = train[[i]], family = binomial(link = "probit"))
print(probit_[[i]])
}

#Naive LPM

LPM_<-list()
for (i in seq(1,length(train))){
LPM_[[i]]=lm(inlf~kids + age + educ + nwifeinc + unem + city, data = train[[i]])
print(LPM_[[i]])
}



```

Once we have computed the naive First Step regression (notice that we will only care about prediction) we will make use of the machine learners in order to predict the participation of women in our sample. We will also exploit the HD created by interactions and powers of the covariates in our sample. For that, we will use a set of learners that will be trained in the i-th training set and tested in the i-th test set, where $i \in \{1,..,5\}$:

```{r 8, include=FALSE, echo=TRUE}
#### Model Fitting


###Elastic Net Regression

train_HD<-list()
test_HD<-list()

for (i in seq(1,5)){


 train_HD[[i]]  <- train[[i]][c(-2:-4,-7:-8,-21,-23)]
 test_HD[[i]]  <- test[[i]][c(-2:-4,-7:-8,-21,-23)]
}
 

y<-list()
x<-list()

for (i in seq(1,5)){
  
x[[i]] <- data.matrix(train_HD[[i]][c(-1)])
y[[i]] <- data.matrix(train_HD[[i]][c(1)])

} 
##Lasso


#Logit Lasso

cv.logit_lasso<-list()

for (i in seq(1,5)){
cv.logit_lasso[[i]]<- cv.glmnet(x[[i]],y[[i]], alpha = 1, family = "binomial")
}


best_lambda_logit_lasso<-list()

for (i in seq(1,5)){
best_lambda_logit_lasso[[i]]<-cv.logit_lasso[[i]]$lambda.min
}


logit_lasso<-list()

for (i in seq(1,5)){

logit_lasso[[i]]<-glmnet(x[[i]],y[[i]], alpha = 1, family = "binomial",lambda = best_lambda_logit_lasso[[i]])
print(logit_lasso[[i]])

}

#Probit Lasso

cv.probit_lasso<-list()

for (i in seq(1,5)){
  cv.probit_lasso[[i]]<- cv.glmnet(x[[i]],y[[i]], alpha = 1, family = binomial(link = "probit"))
}


best_lambda_probit_lasso<-list()

for (i in seq(1,5)){
  best_lambda_probit_lasso[[i]]<-cv.probit_lasso[[i]]$lambda.min
}


probit_lasso<-list()

for (i in seq(1,5)){
  
  probit_lasso[[i]]<-glmnet(x[[i]],y[[i]], alpha = 1, family = binomial(link = "probit"),lambda = best_lambda_probit_lasso[[i]])
  print(probit_lasso[[i]])
  
}

#LPM Lasso 

cv.LPM_lasso<-list()

for (i in seq(1,5)){
cv.LPM_lasso[[i]]<- cv.glmnet(x[[i]],y[[i]], alpha = 1)
}


best_lambda_LPM_lasso<-list()

for (i in seq(1,5)){
best_lambda_LPM_lasso[[i]]<-cv.LPM_lasso[[i]]$lambda.min
}


LPM_lasso<-list()

for (i in seq(1,5)){

LPM_lasso[[i]]<-glmnet(x[[i]],y[[i]], alpha = 1,lambda = best_lambda_LPM_lasso[[i]])
print(LPM_lasso[[i]])

}

##Ridge


#Logit Ridge

cv.logit_ridge<-list()

for (i in seq(1,5)){
cv.logit_ridge[[i]]<- cv.glmnet(x[[i]],y[[i]], alpha = 0, family = "binomial")
}


best_lambda_logit_ridge<-list()

for (i in seq(1,5)){
best_lambda_logit_ridge[[i]]<-cv.logit_ridge[[i]]$lambda.min
}


logit_ridge<-list()

for (i in seq(1,5)){

logit_ridge[[i]]<-glmnet(x[[i]],y[[i]], alpha = 0, family = "binomial",lambda = best_lambda_logit_ridge[[i]])
print(logit_ridge[[i]])

}



#Probit Ridge

cv.probit_ridge<-list()

for (i in seq(1,5)){
cv.probit_ridge[[i]]<- cv.glmnet(x[[i]],y[[i]], alpha = 0, family = binomial(link = "probit"))
}


best_lambda_probit_ridge<-list()

for (i in seq(1,5)){
best_lambda_probit_ridge[[i]]<-cv.probit_ridge[[i]]$lambda.min
}


probit_ridge<-list()

for (i in seq(1,5)){

probit_ridge[[i]]<-glmnet(x[[i]],y[[i]], alpha = 0, family = binomial(link = "probit"),lambda = best_lambda_probit_ridge[[i]])
print(probit_ridge[[i]])

}


#LPM Ridge 

cv.LPM_ridge<-list()

for (i in seq(1,5)){
cv.LPM_ridge[[i]]<- cv.glmnet(x[[i]],y[[i]], alpha = 0)
}


best_lambda_LPM_ridge<-list()

for (i in seq(1,5)){
best_lambda_LPM_ridge[[i]]<-cv.LPM_ridge[[i]]$lambda.min
}


LPM_ridge<-list()

for (i in seq(1,5)){

LPM_ridge[[i]]<-glmnet(x[[i]],y[[i]], alpha = 0,lambda = best_lambda_LPM_ridge[[i]])
print(LPM_ridge[[i]])

}


###Discriminant Analysis

## Linear Discriminant Analysis (LDA)

LDA<-list()

for (i in seq(1,5)){

  LDA[[i]]<-lda(inlf~.,data=train_HD[[i]])
}

 for (i in seq(1,5)){
  plot(LDA[[i]])
   } 

###Classification Trees

##Decision Tree

Decision_Tree<-list()

for (i in seq(1,5)){
  
  Decision_Tree[[i]]<-rpart(inlf~.,data=train_HD[[i]],method = "class")

}


##Random Forest

y_as.factor<-list()

for (i in seq(1,5)){
y_as.factor[[i]]=as.factor(y[[i]])
  }

Random_Forest<-list()

for (i in seq(1,5)){
  
  Random_Forest[[i]]<-randomForest(x=x[[i]],y=y_as.factor[[i]],method = "class")

}

##K-Nearest Neighbors (kNN)


 
kNN.24<-list()
kNN.25<-list()
table_24<-list()
table_25<-list()

for (i in seq(1,5)){
train_kNN<-train_HD[[i]]
test_kNN<-test_HD[[i]]
y_train_kNN<-ifelse(train_kNN[,1]==0,"0", "1")
y_test_kNN<-ifelse(test_kNN[,1]==0,"0", "1")
train_kNN<-train_kNN[,-1]
test_kNN<-test_kNN[,-1]
kNN.24[[i]]<-knn(train = train_kNN, test = test_kNN, cl=y_train_kNN, k=24)
kNN.25[[i]]<-knn(train = train_kNN, test = test_kNN, cl=y_train_kNN, k=25)
table_24[[i]]<-table(kNN.24[[i]], y_test_kNN)
mean(kNN.24[[i]], y_test_kNN)
table_25[[i]]<-table(kNN.25[[i]], y_test_kNN)
mean(kNN.25[[i]], y_test_kNN)
}

for (i in seq(1,5)){
  plot(cv.logit_lasso[[i]])
  plot(cv.LPM_lasso[[i]])
  plot(cv.logit_ridge[[i]])
  plot(cv.probit_ridge[[i]])
  plot(cv.LPM_ridge[[i]])
}


for (i in seq(1,5)){

  rpart.plot( Decision_Tree[[i]],type = 3, box.palette = c("red", "green"), fallen.leaves = TRUE)
  plotcp(Decision_Tree[[i]])
  }

```

```{r 9, include=TRUE, echo=TRUE}

  plot(cv.logit_lasso[[1]])
  plot(cv.LPM_lasso[[1]])
  plot(cv.logit_ridge[[1]])
  plot(cv.probit_ridge[[1]])
  plot(cv.LPM_ridge[[1]])





  rpart.plot( Decision_Tree[[1]],type = 3, box.palette = c("red", "green"), fallen.leaves = TRUE)
  plotcp(Decision_Tree[[1]])
  

```

In this plots (notice that we only plot the first splits) we can see that taking into account $\lambda=0\Rightarrow log(\lambda)=\infty$ that even though we do not have $p>n$ given our number of covariates using Lasso or Ridge may have some justification as we do not archive the minimum for the measure of performing for lambdas equal to 0. Take into account that this will vary across our 5 splits of the training sample. We will display also some measures of performance of the Decision Tree learner and the a graph of this Decision Tree for the 5 splits.

```{r 10, include=FALSE, echo=TRUE}

matrix_RCC<-matrix(NA, nrow=6, ncol=14)
colnames(matrix_RCC)<-c("Naive Logit","Naive Probit","Naive LPM","Logit Lasso","Probit Lasso","LPM Lasso","Logit Ridge", "Probit Ridge","LPM Ridge","LDA" , "Decision Tree" , "Random Forest", "K-Nearest Neighbors (K=24)", "K-Nearest Neighbors (K=25)")
rownames(matrix_RCC)<-c("Split 1", "Split 2", "Split 3", "Split 4", "Split 5", "Average")
matrix_RCC <- rownames_to_column(as.data.frame(matrix_RCC), var = "Rate of Correct Classification (RCC)")

# Note: RCC is Rate of Correct Classification

y_test<-list()

for (i in seq(1,5)){

  y_test[[i]] <- data.matrix(test[[i]][c(1)])
}


y_test_HD<-list()
x_test_HD<-list()

for (i in seq(1,5)){

  y_test_HD[[i]] <- data.matrix(test_HD[[i]][c(1)])
  x_test_HD[[i]] <- data.matrix(test_HD[[i]][c(-1)])
  }

y_as.factor_test<-list()

for (i in seq(1,5)){
y_as.factor_test[[i]]=as.factor(y_test_HD[[i]])
  }

#### Naive Model Testing


test.probs.logit_<-list()
pred.logit_<-list()
RCC.logit_<-list()

test.probs.probit_<-list()
pred.probit_<-list()
RCC.probit_<-list()

test.probs.LPM_<-list()
pred.LPM_<-list()
RCC.LPM_<-list()

test.probs.logit_lasso<-list()
pred.logit_lasso<-list()
RCC.logit_lasso<-list()

test.probs.probit_lasso<-list()
pred.probit_lasso<-list()
RCC.probit_lasso<-list()


test.probs.LPM_lasso<-list()
pred.LPM_lasso<-list()
RCC.LPM_lasso<-list()

test.probs.logit_ridge<-list()
pred.logit_ridge<-list()
RCC.logit_ridge<-list()


test.probs.probit_ridge<-list()
pred.probit_ridge<-list()
RCC.probit_ridge<-list()


test.probs.LPM_ridge<-list()
pred.LPM_ridge<-list()
RCC.LPM_ridge<-list()

test.probs.LDA<-list()
RCC.LDA<-list()


test.probs.Decision_Tree<-list()
RCC.Decision_Tree<-list()

test.probs.Random_Forest<-list()
RCC.Random_Forest<-list()


RCC.kNN.24<-list()

RCC.kNN.25<-list()

for (i in seq(1,5)){
 
#Naive Logit
test.probs.logit_[[i]]<-predict(logit_[[i]], newdata =test[[i]],type="response")
pred.logit_[[i]]<-ifelse(test.probs.logit_[[i]]>0.5, 1, 0)
table(pred.logit_[[i]],y_test[[i]])
RCC.logit_[[i]]<-mean(pred.logit_[[i]]==y_test[[i]])
matrix_RCC[i,2]<-RCC.logit_[[i]]

#Naive Probit
test.probs.probit_[[i]]<-predict(probit_[[i]],newdata =test[[i]],type="response")
pred.probit_[[i]]<-ifelse(test.probs.probit_[[i]]>0.5, 1, 0)
table(pred.probit_[[i]],y_test[[i]])
RCC.probit_[[i]]<-mean(pred.probit_[[i]]==y_test[[i]])
matrix_RCC[i,3]<-RCC.probit_[[i]]


#Naive LPM


test.probs.LPM_[[i]]<-predict(LPM_[[i]],newdata =test[[i]],type="response")
pred.LPM_[[i]]<-ifelse(test.probs.LPM_[[i]]>0.5, 1, 0)
table(pred.LPM_[[i]],y_test[[i]])
RCC.LPM_[[i]]<-mean(pred.LPM_[[i]]==y_test[[i]])
matrix_RCC[i,4]<-RCC.LPM_[[i]]

#### HD Model Testing

# Logit Lasso

test.probs.logit_lasso[[i]]<- logit_lasso[[i]] %>% predict(newx = x_test_HD[[i]])
pred.logit_lasso[[i]]<-ifelse(test.probs.logit_lasso[[i]]>0.5, 1, 0)
table(pred.logit_lasso[[i]],y_test_HD[[i]])
RCC.logit_lasso[[i]]<-mean(pred.logit_lasso[[i]]==y_test_HD[[i]])
matrix_RCC[i,5]<-RCC.logit_lasso[[i]]

# Probit Lasso

test.probs.probit_lasso[[i]]<- probit_lasso[[i]] %>% predict(newx = x_test_HD[[i]])
pred.probit_lasso[[i]]<-ifelse(test.probs.probit_lasso[[i]]>0.5, 1, 0)
table(pred.probit_lasso[[i]],y_test_HD[[i]])
RCC.probit_lasso[[i]]<-mean(pred.probit_lasso[[i]]==y_test_HD[[i]])
matrix_RCC[i,6]<-RCC.probit_lasso[[i]]


# LPM Lasso

test.probs.LPM_lasso[[i]]<- LPM_lasso[[i]] %>% predict(newx = x_test_HD[[i]])
pred.LPM_lasso[[i]]<-ifelse(test.probs.LPM_lasso[[i]]>0.5, 1, 0)
table(pred.LPM_lasso[[i]],y_test_HD[[i]])
RCC.LPM_lasso[[i]]<-mean(pred.LPM_lasso[[i]]==y_test_HD[[i]])
matrix_RCC[i,7]<-RCC.LPM_lasso[[i]]


# Logit Ridge

test.probs.logit_ridge[[i]]<- logit_ridge[[i]] %>% predict(newx = x_test_HD[[i]])
pred.logit_ridge[[i]]<-ifelse(test.probs.logit_ridge[[i]]>0.5, 1, 0)
table(pred.logit_ridge[[i]],y_test_HD[[i]])
RCC.logit_ridge[[i]]<-mean(pred.logit_ridge[[i]]==y_test_HD[[i]])
matrix_RCC[i,8]<-RCC.logit_ridge[[i]]


# Probit Ridge

test.probs.probit_ridge[[i]]<- probit_ridge[[i]] %>% predict(newx = x_test_HD[[i]])
pred.probit_ridge[[i]]<-ifelse(test.probs.probit_ridge[[i]]>0.5, 1, 0)
table(pred.probit_ridge[[i]],y_test_HD[[i]])
RCC.probit_ridge[[i]]<-mean(pred.probit_ridge[[i]]==y_test_HD[[i]])
matrix_RCC[i,9]<-RCC.probit_ridge[[i]]

# LPM Ridge

test.probs.LPM_ridge[[i]]<- LPM_ridge[[i]] %>% predict(newx = x_test_HD[[i]])
pred.LPM_ridge[[i]]<-ifelse(test.probs.LPM_ridge[[i]]>0.5, 1, 0)
table(pred.LPM_ridge[[i]],y_test_HD[[i]])
RCC.LPM_ridge[[i]]<-mean(pred.LPM_ridge[[i]]==y_test_HD[[i]])
matrix_RCC[i,10]<-RCC.LPM_ridge[[i]]

#LDA

test.probs.LDA[[i]]<-predict(LDA[[i]],newdata =test_HD[[i]])
table(test.probs.LDA[[i]]$class,y_as.factor_test[[i]])
RCC.LDA[[i]]<-mean(test.probs.LDA[[i]]$class==y_as.factor_test[[i]])
matrix_RCC[i,11]<-RCC.LDA[[i]]

###Classification Trees

##Decision Tree

test.probs.Decision_Tree[[i]]<-predict(Decision_Tree[[i]], test_HD[[i]], type = "class")
table(test.probs.Decision_Tree[[i]], y_as.factor_test[[i]])
RCC.Decision_Tree[[i]]<-mean(test.probs.Decision_Tree[[i]]==y_as.factor_test[[i]])
matrix_RCC[i,12]<-RCC.Decision_Tree[[i]]


##Random Forest

test.probs.Random_Forest[[i]]<-predict(Random_Forest[[i]], test_HD[[i]], type = "class")
table(test.probs.Random_Forest[[i]], y_as.factor_test[[i]])
RCC.Random_Forest[[i]]<-mean(test.probs.Random_Forest[[i]]==y_as.factor_test[[i]])
matrix_RCC[i,13]<-RCC.Random_Forest[[i]]


#K-Nearest Neighbors (K=24)

table(kNN.24[[i]],y_as.factor_test[[i]])
RCC.kNN.24[[i]]<-mean(kNN.24[[i]]==y_as.factor_test[[i]])
matrix_RCC[i,14]<-RCC.kNN.24[[i]]

#K-Nearest Neighbors (K=25)

table(kNN.25[[i]],y_as.factor_test[[i]])
RCC.kNN.25[[i]]<-mean(kNN.25[[i]]==y_as.factor_test[[i]])
matrix_RCC[i,15]<-RCC.kNN.25[[i]]

}

mean.RCC.logit_<-mean(unlist(RCC.logit_))
matrix_RCC[6,2]<-mean.RCC.logit_
mean.RCC.probit_<-mean(unlist(RCC.probit_))
matrix_RCC[6,3]<-mean.RCC.probit_
mean.RCC.LPM_<-mean(unlist(RCC.LPM_))
matrix_RCC[6,4]<-mean.RCC.LPM_
mean.RCC.logit_lasso<-mean(unlist(RCC.logit_lasso))
matrix_RCC[6,5]<-mean.RCC.logit_lasso
mean.RCC.probit_lasso<-mean(unlist(RCC.probit_lasso))
matrix_RCC[6,6]<-mean.RCC.probit_lasso
mean.RCC.LPM_lasso<-mean(unlist(RCC.LPM_lasso))
matrix_RCC[6,7]<-mean.RCC.LPM_lasso
mean.RCC.logit_ridge<-mean(unlist(RCC.logit_ridge))
matrix_RCC[6,8]<-mean.RCC.logit_ridge
mean.RCC.probit_ridge<-mean(unlist(RCC.probit_ridge))
matrix_RCC[6,9]<-mean.RCC.probit_ridge
mean.RCC.LPM_ridge<-mean(unlist(RCC.LPM_ridge))
matrix_RCC[6,10]<-mean.RCC.LPM_ridge
mean.RCC.LDA<-mean(unlist(RCC.LDA))
matrix_RCC[6,11]<-mean.RCC.LDA
mean.RCC.Decision_Tree<-mean(unlist(RCC.Decision_Tree))
matrix_RCC[6,12]<-mean.RCC.Decision_Tree
mean.RCC.Random_Forest<-mean(unlist(RCC.Random_Forest))
matrix_RCC[6,13]<-mean.RCC.Random_Forest
mean.RCC.kNN.24<-mean(unlist(RCC.kNN.24))
matrix_RCC[6,14]<-mean.RCC.kNN.24
mean.RCC.kNN.25<-mean(unlist(RCC.kNN.25))
matrix_RCC[6,15]<-mean.RCC.kNN.25


```

```{r 11, include=TRUE, echo=TRUE}
knitr::kable(t(matrix_RCC), align="c", booktabs=T, caption="RCC")%>%kable_styling(latex_option=c("hold_position"))

```

In the previous matrix we can see the Rate of Correct Classification (RCC) for the five splits of the test set and for the average value, that will be the measure of out-of-sample performing.

```{r 12, include=FALSE, echo=TRUE}

confusion_logit_<-list()
confusion_probit_<-list()
confusion_LPM_<-list()
confusion_logit_lasso<-list()
confusion_probit_lasso<-list()
confusion_LPM_lasso<-list()
confusion_logit_ridge<-list()
confusion_probit_ridge<-list()
confusion_LPM_ridge<-list()
confusion_LDA<-list()
confusion_Decision_Tree<-list()
confusion_Random_Forest<-list()
confusion_kNN.24<-list()
confusion_kNN.25<-list()


for (i in seq(1,5)){
confusion_logit_[[i]]<-table(pred.logit_[[i]],y_test[[i]])
confusion_probit_[[i]]<-table(pred.probit_[[i]],y_test[[i]])
confusion_LPM_[[i]]<-table(pred.LPM_[[i]],y_test[[i]])
confusion_logit_lasso[[i]]<-table(pred.logit_lasso[[i]],y_test_HD[[i]])
confusion_probit_lasso[[i]]<-table(pred.probit_lasso[[i]],y_test_HD[[i]])
confusion_LPM_lasso[[i]]<-table(pred.LPM_lasso[[i]],y_test_HD[[i]])
confusion_logit_ridge[[i]]<-table(pred.logit_ridge[[i]],y_test_HD[[i]])
confusion_probit_ridge[[i]]<-table(pred.probit_ridge[[i]],y_test_HD[[i]])
confusion_LPM_ridge[[i]]<-table(pred.LPM_ridge[[i]],y_test_HD[[i]])
confusion_LDA[[i]]<-table(test.probs.LDA[[i]]$class,y_as.factor_test[[i]])
confusion_Decision_Tree[[i]]<-table(test.probs.Decision_Tree[[i]], y_as.factor_test[[i]])
confusion_Random_Forest[[i]]<-table(test.probs.Random_Forest[[i]], y_as.factor_test[[i]])
confusion_kNN.24[[i]]<-table(kNN.24[[i]],y_as.factor_test[[i]])
confusion_kNN.25[[i]]<-table(kNN.25[[i]],y_as.factor_test[[i]])
}

for (i in seq(1,5)){
  print(confusion_logit_lasso[[i]])
  print(confusion_probit_lasso[[i]])
}
```

```{r 13, include=TRUE, echo=TRUE}

  print(confusion_logit_lasso[[1]])
  print(confusion_probit_lasso[[1]])

```

As we can see here the confusion matrix (notice that we only plot the first splits) for the Logit Lasso and the Probit Lasso tell us something similar to what the RCC is telling us. Hence, we will use the measure of RCC as a measure of performing out-of-sample since this measure is consistent with the confusion matrix.

Once we have tested all our learners, by computing the Rate of Correct Classification (RCC) and also the confusion matrices of each split, we will proceed by selecting two different models for our estimation of this First Stage. These model are the Logit Lasso and the Probit Lasso, as they report a RCC of 0.95 and 0.93, respectively. So, once we have selected the models we will use these predictions to compute the Second Stage. In order to do show, we will compute a similar procedure to the one following by Mroz (1987) in the known Mroz's multistage procedure to estimate a structural labor supply equation. This procedure is as follows, we will have the following equations, where the setting is as stated in Weerasooriya (2018), that in our setting is the following one:

$D_i:= 1\{inlf_i=1\}=F(\textbf{Z}_i^{T}\gamma) + e_{i}$ $(1)$

$log(Wage_{i})=\textbf{X}_{1i}^{T}\beta + v_{1i}$ $(2)$

$h_i=\textbf{X}_{2i}^{T}\beta + v_{2i}$ $(3)$

where equation (1) is the Selection equation in the model and equation (2) is the Second Stage. Hence, both equations (1) and (2) accounts for the Heckman two-step procedure. We will have that $\textbf{Z}_i$ in Mroz (1987) are the following covariates, for two different specifications:

$(i):$ family composition (children), powers of woman's age and education and their interactions between them to approximate a third-order polynomial in these two variables,education of the woman's father and mother, local unemployment rate, binary variable on living in a large city, and family income excl. woman's earnings.

$(ii):$ the variables in $(i)$ plus labor market experience and its square.

Whereas in Mroz (1987) this is the set of covariates used in equation (1), we will follow a more data driven way and thus, make use of the HD setting to predict by using machine learner. The two best learners are Logit Lasso and the Probit Lasso. We will compute the Inverse Mill's ratio to account for the sample Selection in our Second Stage. Once we have computed an estimate of $\lambda(\textbf{Z}_i^{T}\gamma)$, say $\hat\lambda(\textbf{Z}_i^{T}\hat\gamma)$, where we can apply the Heckman two-step approach assuming joint normality of $(e_{i},v_{1i}) \sim \mathcal{N_2}(0,,\Sigma)$ or a non-parametric version where we compute $\hat\lambda(\textbf{Z}_i^{T}\hat\gamma)=\dfrac{f_n(\textbf{Z}_i^{T}\hat\gamma)}{F_n(\textbf{Z}_i^{T}\hat\gamma)}$ where $f_n(\textbf{Z}_i^{T}\hat\gamma)$ is the empirical pdf from the predicted values of the machine learner estimate, while $F_n(\textbf{Z}_i^{T}\hat\gamma)$ is the empirical cdf from the predicted values of the machine learner estimate. Once we have compute our estimate of the Inverse Mill's ratio, we will use it in equation (2) where $\textbf{X}_{1i}=(\textbf{Z}_i,\hat\lambda(\textbf{Z}_i^{T}\hat\gamma))$ and by that we will account for the selection bias.

Once, we have solve this selection issue in equation (2) we will proceed by predicting the log(Wage) using in this step also some learners.

Finally, the contents of $\textbf{X}_{2i}$ will be $\hat log(Wage)_i$ from equation (2), the non-wife income, number of older and younger children, the age and education of the wife and $\hat\lambda(\textbf{Z}_i^{T}\hat\gamma)$ from equation (1).

## Second Step estimation

In this step we will proceed by using the estimates of the probit lasso model and the logit lasso model, as these two learners are the one that predicts the better out-of-sample. Given that, we will use both learners to compute the empirical lambda of Heckman we will use for the following formula for computing the empirical cdf:

$$
F_{\textbf{Z}i^{T}\hat\gamma}(b)={\dfrac{1}{n}\sum_{i=1}^{n}} 1 \{ \textbf{Z}_i^{T}\hat\gamma \leqslant b \}
$$

and using the Kernel Density Estimation in order to estimate the empirical pdf. Once, we have compute both empirical functions the empirical lambda de Heckman is just the ratio between both functions.

```{r 14, include=FALSE, echo=TRUE}

x_data_frame<- as.matrix(data_frame[c(-1:-4,-7:-8,-21,-23)])
y_data_frame<-as.matrix(data_frame[c(1)])

#Logit Lasso (full sample (N=753))

cv.logit_lasso_data_frame<- cv.glmnet(x_data_frame,data_frame$inlf, alpha = 1, family = "binomial")
best_lambda_logit_lasso_data_frame<-cv.logit_lasso_data_frame$lambda.min
logit_lasso_data_frame<-glmnet(x_data_frame,y_data_frame, alpha = 1,family = "binomial" ,lambda = best_lambda_logit_lasso_data_frame)

#Probit Lasso (full sample (N=753))

cv.probit_lasso_data_frame<- cv.glmnet(x_data_frame,data_frame$inlf, alpha = 1, family = binomial(link = "probit"))
best_lambda_probit_lasso_data_frame<-cv.probit_lasso_data_frame$lambda.min
probit_lasso_data_frame<-glmnet(x_data_frame,y_data_frame, alpha = 1, family = binomial(link = "probit"),lambda = best_lambda_probit_lasso_data_frame)

##Compute the non-parametric lambda of Heckman 

x_beta_hat_logit_lasso<-predict(logit_lasso_data_frame, newx=data.matrix(data_frame[c(-1:-4,-7:-8,-21,-23 )]))
x_beta_hat_probit_lasso<-predict(probit_lasso_data_frame, newx=data.matrix(data_frame[c(-1:-4,-7:-8,-21,-23)]))


#Empirical cdf for the predicted values

ecdf_x_beta_hat_logit_lasso<-ecdf(x_beta_hat_logit_lasso)
data_frame$ecdf_logit_lasso<-ecdf_x_beta_hat_logit_lasso(x_beta_hat_logit_lasso)  
ecdf_x_beta_hat_probit_lasso<-ecdf(x_beta_hat_probit_lasso)
data_frame$ecdf_probit_lasso<-ecdf_x_beta_hat_probit_lasso(x_beta_hat_probit_lasso)




#Empirical pdf for the predicted values

epdf_x_beta_hat_logit_lasso<-density(x_beta_hat_logit_lasso,n = 753)
data_frame$epdf_logit_lasso<-epdf_x_beta_hat_logit_lasso[["y"]]
epdf_x_beta_hat_probit_lasso<-density(x_beta_hat_probit_lasso, n = 753)
data_frame$epdf_probit_lasso<-epdf_x_beta_hat_probit_lasso[["y"]]


#Empirical lambda of Heckman for the predicted values

data_frame$lambda_logit_lasso<-data_frame$epdf_logit_lasso/data_frame$ecdf_logit_lasso
data_frame$lambda_probit_lasso<-data_frame$epdf_probit_lasso/data_frame$ecdf_probit_lasso

plot(1:30)


```

```{r 15, include=TRUE, echo=TRUE}

plot(ecdf_x_beta_hat_logit_lasso,main = "Emprirical cdf from the predicted values of the Logit Lasso estimates")
plot(ecdf_x_beta_hat_probit_lasso, main = "Emprirical cdf from the predicted values of the Probit Lasso estimates")


plot(epdf_x_beta_hat_logit_lasso, main = "Emprirical pdf from the predicted values of the Logit Lasso estimates")
plot(epdf_x_beta_hat_probit_lasso,  main = "Emprirical pdf from the predicted values of the Probit Lasso estimates")

plot(data_frame$lambda_logit_lasso, main = "Emprirical lambda of Heckman from the predicted values of the Logit Lasso estimates")
plot(data_frame$lambda_probit_lasso, main = "Emprirical lambda of Heckman from the predicted values of the Logit Lasso estimates")

```

Now, we will proceed by training some learners in order to predict log(Wage) in equation (2) and in order to do so accounting for sample selection we will use the empirical lambda of Heckman. Hence, as we proceed by computing two versions of the empirical lambda of Heckman, from this point until the estimation of the structural labour supply equation for women we will have two different versions of the predicted log(Wage) and two versions of the the structural labour supply equation for women, say two versions of equation (3).

In the following step we will use the five divisions that we already compute in previous steps and use the two train our learners and in the following step test them. Take into account that previously we have equation (1) where the outcome is a binary variable, which implies that we use as a measure to test the learners the RCC, as stated in James, Witten, Hastie, & Tibshirani (2013). But in equation (2) we would not have classification, as our outcome is a continuous variable, then our measure in order to test the learners is going to be the MSE.

```{r 16, include=FALSE, echo=TRUE}

participants_data_frame<-subset(data_frame, inlf == 1)

sample_participants<-list()
train_participants<-list()
test_participants<-list()

for (i in seq(1,5)){

 sample_participants[[i]]<-sample(c(TRUE, FALSE), nrow(participants_data_frame), replace=TRUE, prob=c(0.8,0.2))
 train_participants[[i]]  <- participants_data_frame[sample_participants[[i]], ]
 test_participants[[i]]  <- participants_data_frame[!sample_participants[[i]], ]
}
  

train_2Step_lambda_logit<-list()
test_2Step_lambda_logit<-list()
train_2Step_lambda_probit<-list()
test_2Step_lambda_probit<-list()

for (i in seq(1,5)){


 train_2Step_lambda_logit[[i]]  <- train_participants[[i]][c(-2:-4,-7:-8,-23,-172:-175,-177)]
 test_2Step_lambda_logit[[i]]  <- test_participants[[i]][c(-2:-4,-7:-8,-23,-172:-175,-177)]
 train_2Step_lambda_probit[[i]]  <- train_participants[[i]][c(-2:-4,-7:-8,-23,-172:-176)]
 test_2Step_lambda_probit[[i]]  <- test_participants[[i]][c(-2:-4,-7:-8,-23,-172:-176)]

 
 }
 

y_lambda_logit<-list()
x_lambda_logit<-list()
y_lambda_probit<-list()
x_lambda_probit<-list()

for (i in seq(1,5)){
  
x_lambda_logit[[i]] <- data.matrix(train_2Step_lambda_logit[[i]][c(-16)])
y_lambda_logit[[i]] <- data.matrix(train_2Step_lambda_logit[[i]][c(16)])
x_lambda_probit[[i]] <- data.matrix(train_2Step_lambda_probit[[i]][c(-16)])
y_lambda_probit[[i]] <- data.matrix(train_2Step_lambda_probit[[i]][c(16)])

} 


x_lambda_logit_test<-list()
x_lambda_probit_test<-list()

for (i in seq(1,5)){
  
x_lambda_logit_test[[i]] <- data.matrix(test_2Step_lambda_logit[[i]][c(-16)])
x_lambda_probit_test[[i]] <- data.matrix(test_2Step_lambda_probit[[i]][c(-16)])

} 

#### Model Fitting (for empirical lambda of Heckman for Logit Lasso)

###OLS

OLS_lambda_logit<-list()
for (i in seq(1,5)){
OLS_lambda_logit[[i]]=lm(lwage~., data = train_2Step_lambda_logit[[i]])
}


###Elastic Net Regression

##Lasso

#OLS Lasso 

cv.OLS_lasso_lambda_logit<-list()

for (i in seq(1,5)){
cv.OLS_lasso_lambda_logit[[i]]<- cv.glmnet(x_lambda_logit[[i]],y_lambda_logit[[i]], alpha = 1)
}


best_lambda_OLS_lasso_lambda_logit<-list()

for (i in seq(1,5)){
best_lambda_OLS_lasso_lambda_logit[[i]]<-cv.OLS_lasso_lambda_logit[[i]]$lambda.min
}


OLS_lasso_lambda_logit<-list()

for (i in seq(1,5)){

OLS_lasso_lambda_logit[[i]]<-glmnet(x_lambda_logit[[i]],y_lambda_logit[[i]], alpha = 1,lambda = best_lambda_OLS_lasso_lambda_logit[[i]])
print(OLS_lasso_lambda_logit[[i]])

}

##Ridge

#OLS Ridge 

cv.OLS_ridge_lambda_logit<-list()

for (i in seq(1,5)){
cv.OLS_ridge_lambda_logit[[i]]<- cv.glmnet(x_lambda_logit[[i]],y_lambda_logit[[i]], alpha = 0)
}


best_lambda_OLS_ridge_lambda_logit<-list()

for (i in seq(1,5)){
best_lambda_OLS_ridge_lambda_logit[[i]]<-cv.OLS_ridge_lambda_logit[[i]]$lambda.min
}


OLS_ridge_lambda_logit<-list()

for (i in seq(1,5)){

OLS_ridge_lambda_logit[[i]]<-glmnet(x_lambda_logit[[i]],y_lambda_logit[[i]], alpha = 0,lambda = best_lambda_OLS_ridge_lambda_logit[[i]])
print(OLS_ridge_lambda_logit[[i]])

}


# alpha=0.5


cv.OLS_EN_lambda_logit<-list()

for (i in seq(1,5)){
cv.OLS_EN_lambda_logit[[i]]<- cv.glmnet(x_lambda_logit[[i]],y_lambda_logit[[i]], alpha = 0.5)
}


best_lambda_OLS_EN_lambda_logit<-list()

for (i in seq(1,5)){
best_lambda_OLS_EN_lambda_logit[[i]]<-cv.OLS_EN_lambda_logit[[i]]$lambda.min
}


OLS_EN_lambda_logit<-list()

for (i in seq(1,5)){

OLS_EN_lambda_logit[[i]]<-glmnet(x_lambda_logit[[i]],y_lambda_logit[[i]], alpha = 0.5,lambda = best_lambda_OLS_EN_lambda_logit[[i]])
print(OLS_EN_lambda_logit[[i]])

}


###Classification Trees

##Decision Tree

Decision_Tree_lambda_logit<-list()

for (i in seq(1,5)){
  
  Decision_Tree_lambda_logit[[i]]<-rpart(lwage~.,data=train_2Step_lambda_logit[[i]])

}



##Random Forest


Random_Forest_lambda_logit<-list()

for (i in seq(1,5)){
  
  Random_Forest_lambda_logit[[i]]<-randomForest(lwage~.,data=train_2Step_lambda_logit[[i]])

}

##Neural Networks (NN)

NN_lambda_logit<-list()

for (i in seq(1,5)){
  
  NN_lambda_logit[[i]]<-neuralnet(lwage~.,data=train_2Step_lambda_logit[[i]], hidden = c(2,2))

}


#### Model Fitting (for empirical lambda of Heckman for Probit Lasso)

###OLS

OLS_lambda_probit<-list()
for (i in seq(1,5)){
OLS_lambda_probit[[i]]=lm(lwage~., data = train_2Step_lambda_probit[[i]])
}


###Elastic Net Regression

##Lasso

#OLS Lasso 

cv.OLS_lasso_lambda_probit<-list()

for (i in seq(1,5)){
cv.OLS_lasso_lambda_probit[[i]]<- cv.glmnet(x_lambda_probit[[i]],y_lambda_probit[[i]], alpha = 1)
}


best_lambda_OLS_lasso_lambda_probit<-list()

for (i in seq(1,5)){
best_lambda_OLS_lasso_lambda_probit[[i]]<-cv.OLS_lasso_lambda_probit[[i]]$lambda.min
}


OLS_lasso_lambda_probit<-list()

for (i in seq(1,5)){

OLS_lasso_lambda_probit[[i]]<-glmnet(x_lambda_probit[[i]],y_lambda_probit[[i]], alpha = 1,lambda = best_lambda_OLS_lasso_lambda_probit[[i]])
print(OLS_lasso_lambda_probit[[i]])

}

##Ridge

#OLS Ridge 

cv.OLS_ridge_lambda_probit<-list()

for (i in seq(1,5)){
cv.OLS_ridge_lambda_probit[[i]]<- cv.glmnet(x_lambda_probit[[i]],y_lambda_probit[[i]], alpha = 0)
}


best_lambda_OLS_ridge_lambda_probit<-list()

for (i in seq(1,5)){
best_lambda_OLS_ridge_lambda_probit[[i]]<-cv.OLS_ridge_lambda_probit[[i]]$lambda.min
}


OLS_ridge_lambda_probit<-list()

for (i in seq(1,5)){

OLS_ridge_lambda_probit[[i]]<-glmnet(x_lambda_probit[[i]],y_lambda_probit[[i]], alpha = 0,lambda = best_lambda_OLS_ridge_lambda_probit[[i]])
print(OLS_ridge_lambda_probit[[i]])

}


# alpha=0.5


cv.OLS_EN_lambda_probit<-list()

for (i in seq(1,5)){
cv.OLS_EN_lambda_probit[[i]]<- cv.glmnet(x_lambda_probit[[i]],y_lambda_probit[[i]], alpha = 0.5)
}


best_lambda_OLS_EN_lambda_probit<-list()

for (i in seq(1,5)){
best_lambda_OLS_EN_lambda_probit[[i]]<-cv.OLS_EN_lambda_probit[[i]]$lambda.min
}


OLS_EN_lambda_probit<-list()

for (i in seq(1,5)){

OLS_EN_lambda_probit[[i]]<-glmnet(x_lambda_probit[[i]],y_lambda_probit[[i]], alpha = 0.5,lambda = best_lambda_OLS_EN_lambda_probit[[i]])
print(OLS_EN_lambda_probit[[i]])

}


###Classification Trees

##Decision Tree

Decision_Tree_lambda_probit<-list()

for (i in seq(1,5)){
  
  Decision_Tree_lambda_probit[[i]]<-rpart(lwage~.,data=train_2Step_lambda_probit[[i]])

}



##Random Forest


Random_Forest_lambda_probit<-list()

for (i in seq(1,5)){
  
  Random_Forest_lambda_probit[[i]]<-randomForest(lwage~.,data=train_2Step_lambda_probit[[i]])

}

##Neural Networks (NN)

NN_lambda_probit<-list()

for (i in seq(1,5)){
  
  NN_lambda_probit[[i]]<-neuralnet(lwage~.,data=train_2Step_lambda_probit[[i]], hidden = c(2,2))

}

for (i in seq(1,5)){

  rpart.plot( Decision_Tree_lambda_logit[[i]],type = 3, box.palette = c("red", "green"), fallen.leaves = TRUE)
  plotcp(Decision_Tree_lambda_logit[[i]])
  }

for (i in seq(1,5)){

  rpart.plot( Decision_Tree_lambda_probit[[i]],type = 3, box.palette = c("red", "green"), fallen.leaves = TRUE)
  plotcp(Decision_Tree_lambda_probit[[i]])
  }

```

```{r 17, include=TRUE, echo=TRUE}


  rpart.plot( Decision_Tree_lambda_logit[[1]],type = 3, box.palette = c("red", "green"), fallen.leaves = TRUE)
  plotcp(Decision_Tree_lambda_logit[[1]])

  rpart.plot( Decision_Tree_lambda_probit[[1]],type = 3, box.palette = c("red", "green"), fallen.leaves = TRUE)
  plotcp(Decision_Tree_lambda_probit[[1]])
  
```

Once we have trained in five splits the following models (even though we will only show the graphs of the first split):

$(i)$ OLS $(ii)$ OLS Lasso $(iii)$ OLS Ridge $(iv)$ OLS with $\alpha=0.5$, which corresponds to the Elastic Net $(v)$ Decision Tree $(vi)$ Random Forest $(vii)$ Neural Networks

We will use Neural Networks to test whether in this second equation performs as in Paliouras & Jessen (1999), even though these authors compute NN for equation (1). The following step is to test the the learners using the MSE as a measure of performing out-of-sample.

```{r 18, include=FALSE, echo=TRUE}
matrix_MSE_logit<-matrix(NA, nrow=6, ncol=7)
colnames(matrix_MSE_logit)<-c("OLS","OLS Lasso", "OLS Ridge","OLS Elastic Net","Decision Tree","Random Forest","Neural Networks")
rownames(matrix_MSE_logit)<-c("Split 1", "Split 2", "Split 3", "Split 4", "Split 5", "Average")
matrix_MSE_logit <- rownames_to_column(as.data.frame(matrix_MSE_logit), var = "Mean Squared Error (MSE) for empirical lambda of Heckman for Logit Lasso")



test.probs.OLS_lambda_logit<-list()
MSE.OLS_lambda_logit<-list()

test.probs.OLS_lasso_lambda_logit<-list()
MSE.OLS_lasso_lambda_logit<-list()

test.probs.OLS_ridge_lambda_logit<-list()
MSE.OLS_ridge_lambda_logit<-list()

test.probs.OLS_EN_lambda_logit<-list()
MSE.OLS_EN_lambda_logit<-list()

test.probs.Decision_Tree_lambda_logit<-list()
MSE.Decision_Tree_lambda_logit<-list()

test.probs.Random_Forest_lambda_logit<-list()
MSE.Random_Forest_lambda_logit<-list()

test.probs.NN_lambda_logit<-list()
MSE.NN_lambda_logit<-list()

for (i in seq(1,5)){


#OLS


test.probs.OLS_lambda_logit[[i]]<-predict(OLS_lambda_logit[[i]],newdata =test_2Step_lambda_logit[[i]], type="response")
MSE.OLS_lambda_logit[[i]]<-mse(test_2Step_lambda_logit[[i]]$lwage,test.probs.OLS_lambda_logit[[i]])
matrix_MSE_logit[i,2]<-MSE.OLS_lambda_logit[[i]]

#### HD Model Testing

# OLS Lasso

test.probs.OLS_lasso_lambda_logit[[i]]<- OLS_lasso_lambda_logit[[i]] %>% predict(newx =x_lambda_logit_test[[i]])
MSE.OLS_lasso_lambda_logit[[i]]<-mse(test_2Step_lambda_logit[[i]]$lwage,test.probs.OLS_lasso_lambda_logit[[i]])
matrix_MSE_logit[i,3]<-MSE.OLS_lasso_lambda_logit[[i]]


# OLS Ridge

test.probs.OLS_ridge_lambda_logit[[i]]<- OLS_ridge_lambda_logit[[i]] %>% predict(newx =x_lambda_logit_test[[i]])
MSE.OLS_ridge_lambda_logit[[i]]<-mse(test_2Step_lambda_logit[[i]]$lwage,test.probs.OLS_ridge_lambda_logit[[i]])
matrix_MSE_logit[i,4]<-MSE.OLS_ridge_lambda_logit[[i]]

# OLS Elastic Nets

test.probs.OLS_EN_lambda_logit[[i]]<- OLS_EN_lambda_logit[[i]] %>% predict(newx =x_lambda_logit_test[[i]])
MSE.OLS_EN_lambda_logit[[i]]<-mse(test_2Step_lambda_logit[[i]]$lwage,test.probs.OLS_EN_lambda_logit[[i]])
matrix_MSE_logit[i,5]<-MSE.OLS_EN_lambda_logit[[i]]


###Classification Trees

##Decision Tree

test.probs.Decision_Tree_lambda_logit[[i]]<-predict(Decision_Tree_lambda_logit[[i]],newdata =test_2Step_lambda_logit[[i]], type="matrix")
MSE.Decision_Tree_lambda_logit[[i]]<-mse(test_2Step_lambda_logit[[i]]$lwage,test.probs.Decision_Tree_lambda_logit[[i]])
matrix_MSE_logit[i,6]<-MSE.Decision_Tree_lambda_logit[[i]]


##Random Forest

test.probs.Random_Forest_lambda_logit[[i]]<-predict(Random_Forest_lambda_logit[[i]], newdata =test_2Step_lambda_logit[[i]], type="response")
MSE.Random_Forest_lambda_logit[[i]]<-mse(test_2Step_lambda_logit[[i]]$lwage,test.probs.Random_Forest_lambda_logit[[i]])
matrix_MSE_logit[i,7]<-MSE.Random_Forest_lambda_logit[[i]]


#Neural Networks

test.probs.NN_lambda_logit[[i]]<-predict(NN_lambda_logit[[i]],  newdata =test_2Step_lambda_logit[[i]], type="response")
MSE.NN_lambda_logit[[i]]<-mse(test_2Step_lambda_logit[[i]]$lwage,test.probs.NN_lambda_logit[[i]])
matrix_MSE_logit[i,8]<-MSE.NN_lambda_logit[[i]]


}

mean.MSE.OLS_lambda_logit<-mean(unlist(MSE.OLS_lambda_logit))
matrix_MSE_logit[6,2]<-mean.MSE.OLS_lambda_logit
mean.MSE.OLS_lasso_lambda_logit<-mean(unlist(MSE.OLS_lasso_lambda_logit))
matrix_MSE_logit[6,3]<-mean.MSE.OLS_lasso_lambda_logit
mean.MSE.OLS_ridge_lambda_logit<-mean(unlist(MSE.OLS_ridge_lambda_logit))
matrix_MSE_logit[6,4]<-mean.MSE.OLS_ridge_lambda_logit
mean.MSE.OLS_EN_lambda_logit<-mean(unlist(MSE.OLS_EN_lambda_logit))
matrix_MSE_logit[6,5]<-mean.MSE.OLS_EN_lambda_logit
mean.MSE.Decision_Tree_lambda_logit<-mean(unlist(MSE.Decision_Tree_lambda_logit))
matrix_MSE_logit[6,6]<-mean.MSE.Decision_Tree_lambda_logit
mean.MSE.Random_Forest_lambda_logit<-mean(unlist(MSE.Random_Forest_lambda_logit))
matrix_MSE_logit[6,7]<-mean.MSE.Random_Forest_lambda_logit
mean.MSE.NN_lambda_logit<-mean(unlist(MSE.NN_lambda_logit))
matrix_MSE_logit[6,8]<-mean.MSE.NN_lambda_logit


```

```{r 19, include=TRUE, echo=TRUE}
knitr::kable(t(matrix_MSE_logit), align="c", booktabs=T, caption="MSE Logit")%>%kable_styling(latex_option=c("hold_position"))
```

Here, for the case in which we use the empirical lambda of Heckman for the predicted values of the Logit Lasso model, we have 3 learners that predict the best out-of-sample. These learners are the OLS Lasso (Average MSE=0.2916950), the OLS EN \(\lambda=0.5\) (Average MSE=0.2988637) and the OLS Ridge (Average MSE=0.3364880). In this case, we will use the OLS Lasso as the learner in order to predict in
the whole sample the value of \(\hat {log(Wage)}\).

```{r 20, include=FALSE, echo=TRUE}
matrix_MSE_probit<-matrix(NA, nrow=6, ncol=7)
colnames(matrix_MSE_probit)<-c("OLS","OLS Lasso", "OLS Ridge","OLS Elastic Net","Decision Tree","Random Forest","Neural Networks")
rownames(matrix_MSE_probit)<-c("Split 1", "Split 2", "Split 3", "Split 4", "Split 5", "Average")
matrix_MSE_probit <- rownames_to_column(as.data.frame(matrix_MSE_probit), var = "Mean Squared Error (MSE) for empirical lambda of Heckman for Probit Lasso")



test.probs.OLS_lambda_probit<-list()
MSE.OLS_lambda_probit<-list()

test.probs.OLS_lasso_lambda_probit<-list()
MSE.OLS_lasso_lambda_probit<-list()

test.probs.OLS_ridge_lambda_probit<-list()
MSE.OLS_ridge_lambda_probit<-list()

test.probs.OLS_EN_lambda_probit<-list()
MSE.OLS_EN_lambda_probit<-list()

test.probs.Decision_Tree_lambda_probit<-list()
MSE.Decision_Tree_lambda_probit<-list()

test.probs.Random_Forest_lambda_probit<-list()
MSE.Random_Forest_lambda_probit<-list()

test.probs.NN_lambda_probit<-list()
MSE.NN_lambda_probit<-list()

for (i in seq(1,5)){


#OLS


test.probs.OLS_lambda_probit[[i]]<-predict(OLS_lambda_probit[[i]],newdata =test_2Step_lambda_probit[[i]], type="response")
MSE.OLS_lambda_probit[[i]]<-mse(test_2Step_lambda_probit[[i]]$lwage,test.probs.OLS_lambda_probit[[i]])
matrix_MSE_probit[i,2]<-MSE.OLS_lambda_probit[[i]]

#### HD Model Testing

# OLS Lasso

test.probs.OLS_lasso_lambda_probit[[i]]<- OLS_lasso_lambda_probit[[i]] %>% predict(newx =x_lambda_probit_test[[i]])
MSE.OLS_lasso_lambda_probit[[i]]<-mse(test_2Step_lambda_probit[[i]]$lwage,test.probs.OLS_lasso_lambda_probit[[i]])
matrix_MSE_probit[i,3]<-MSE.OLS_lasso_lambda_probit[[i]]


# OLS Ridge

test.probs.OLS_ridge_lambda_probit[[i]]<- OLS_ridge_lambda_probit[[i]] %>% predict(newx =x_lambda_probit_test[[i]])
MSE.OLS_ridge_lambda_probit[[i]]<-mse(test_2Step_lambda_probit[[i]]$lwage,test.probs.OLS_ridge_lambda_probit[[i]])
matrix_MSE_probit[i,4]<-MSE.OLS_ridge_lambda_probit[[i]]

# OLS Elastic Nets

test.probs.OLS_EN_lambda_probit[[i]]<- OLS_EN_lambda_probit[[i]] %>% predict(newx =x_lambda_probit_test[[i]])
MSE.OLS_EN_lambda_probit[[i]]<-mse(test_2Step_lambda_probit[[i]]$lwage,test.probs.OLS_EN_lambda_probit[[i]])
matrix_MSE_probit[i,5]<-MSE.OLS_EN_lambda_probit[[i]]


###Classification Trees

##Decision Tree

test.probs.Decision_Tree_lambda_probit[[i]]<-predict(Decision_Tree_lambda_probit[[i]],newdata =test_2Step_lambda_probit[[i]], type="matrix")
MSE.Decision_Tree_lambda_probit[[i]]<-mse(test_2Step_lambda_probit[[i]]$lwage,test.probs.Decision_Tree_lambda_probit[[i]])
matrix_MSE_probit[i,6]<-MSE.Decision_Tree_lambda_probit[[i]]


##Random Forest

test.probs.Random_Forest_lambda_probit[[i]]<-predict(Random_Forest_lambda_probit[[i]], newdata =test_2Step_lambda_probit[[i]], type="response")
MSE.Random_Forest_lambda_probit[[i]]<-mse(test_2Step_lambda_probit[[i]]$lwage,test.probs.Random_Forest_lambda_probit[[i]])
matrix_MSE_probit[i,7]<-MSE.Random_Forest_lambda_probit[[i]]


#Neural Networks

test.probs.NN_lambda_probit[[i]]<-predict(NN_lambda_probit[[i]],  newdata =test_2Step_lambda_probit[[i]], type="response")
MSE.NN_lambda_probit[[i]]<-mse(test_2Step_lambda_probit[[i]]$lwage,test.probs.NN_lambda_probit[[i]])
matrix_MSE_probit[i,8]<-MSE.NN_lambda_probit[[i]]


}

mean.MSE.OLS_lambda_probit<-mean(unlist(MSE.OLS_lambda_probit))
matrix_MSE_probit[6,2]<-mean.MSE.OLS_lambda_probit
mean.MSE.OLS_lasso_lambda_probit<-mean(unlist(MSE.OLS_lasso_lambda_probit))
matrix_MSE_probit[6,3]<-mean.MSE.OLS_lasso_lambda_probit
mean.MSE.OLS_ridge_lambda_probit<-mean(unlist(MSE.OLS_ridge_lambda_probit))
matrix_MSE_probit[6,4]<-mean.MSE.OLS_ridge_lambda_probit
mean.MSE.OLS_EN_lambda_probit<-mean(unlist(MSE.OLS_EN_lambda_probit))
matrix_MSE_probit[6,5]<-mean.MSE.OLS_EN_lambda_probit
mean.MSE.Decision_Tree_lambda_probit<-mean(unlist(MSE.Decision_Tree_lambda_probit))
matrix_MSE_probit[6,6]<-mean.MSE.Decision_Tree_lambda_probit
mean.MSE.Random_Forest_lambda_probit<-mean(unlist(MSE.Random_Forest_lambda_probit))
matrix_MSE_probit[6,7]<-mean.MSE.Random_Forest_lambda_probit
mean.MSE.NN_lambda_probit<-mean(unlist(MSE.NN_lambda_probit))
matrix_MSE_probit[6,8]<-mean.MSE.NN_lambda_probit



```

```{r 21, include=TRUE, echo=TRUE}
knitr::kable(t(matrix_MSE_probit), align="c", booktabs=T, caption="MSE Probit")%>%kable_styling(latex_option=c("hold_position"))
```

In our case different that in Paliouras & Jessen (1999) we do not have the NN learner as the one that outperforms in out-of-sample prediction. Even though, the NN does well, in this case in which we use the empirical lambda of Heckman for the predicted values of the Probit Lasso the learner that best works out-of-sample is the OLS EN \(\lambda=0.5\) (Average MSE=0.3038293) even if in this case OLS Lasso works well in this environment (Average MSE=0.2911794), we will use the version in which \(\lambda=0.5\) to compute \(\hat {log(Wage)}\) for the whole sample, in order to account for robustness in our results depending on the selection of \(\lambda\).

```{r 22, include=FALSE, echo=TRUE}


x_lambda_logit_final<-as.matrix(participants_data_frame[c(-2:-4,-7:-8,-23,-172:-175,-177,-21)])
y_lambda_logit_final<-as.matrix(participants_data_frame[c(21)])
x_lambda_probit_final<-as.matrix(participants_data_frame[c(-2:-4,-7:-8,-23,-172:-176,-21)])
y_lambda_probit_final<-as.matrix(participants_data_frame[c(21)])

##OLS Lasso (fitting the model for the whole sample of participants)


cv.OLS_lasso_lambda_logit_final<- cv.glmnet(x_lambda_logit_final,y_lambda_logit_final, alpha = 1)
best_lambda_OLS_lasso_lambda_logit_final<-cv.OLS_lasso_lambda_logit_final$lambda.min
OLS_lasso_lambda_logit_final<-glmnet(x_lambda_logit_final,y_lambda_logit_final, alpha = 1,lambda = best_lambda_OLS_lasso_lambda_logit_final)
print(OLS_lasso_lambda_logit_final)
participants_data_frame$lwage_hat_1<- OLS_lasso_lambda_logit_final %>% predict(newx =x_lambda_logit_final)



##OLS EN (fitting the model for the whole sample of participants)


cv.OLS_lasso_lambda_probit_final<- cv.glmnet(x_lambda_probit_final,y_lambda_probit_final, alpha = 0.5)
best_lambda_OLS_lasso_lambda_probit_final<-cv.OLS_lasso_lambda_probit_final$lambda.min
OLS_lasso_lambda_probit_final<-glmnet(x_lambda_probit_final,y_lambda_probit_final, alpha = 0.5,lambda = best_lambda_OLS_lasso_lambda_probit_final)
print(OLS_lasso_lambda_probit_final)
participants_data_frame$lwage_hat_2<- OLS_lasso_lambda_probit_final %>% predict(newx =x_lambda_probit_final)







```

After computing this regressions we will use the predicted values of this regression in order to compute the labour supply equation. As stated in Weerasooriya (2018) in its paper Mroz (1987) suspects that log(Wage) will be endogenous in equation (3), the structural equation. Then, the Reduced Form equation, equation (2) for this porpoise will solve the endogeneity of the structural equation (3) and also we will solve the Sample Selection issue by computing introduction as covariates in equation (3) the empirical Inverse Mill's ratio.

```{r 23, include=TRUE, echo=TRUE}
#Structural Labour Supply equations
attach(participants_data_frame)

structural_equation_1<-lm(hours~lwage_hat_1+nwifeinc+age+educ+kids+lambda_logit_lasso )
structural_equation_1_robust<-lmtest::coeftest(structural_equation_1, vcoc=sandwich::vcovHC(structural_equation_2, type='HC2'))
 
structural_equation_2<-lm(hours~lwage_hat_2+nwifeinc+age+educ+kids+lambda_probit_lasso )
structural_equation_2_robust<-lmtest::coeftest(structural_equation_2, vcoc=sandwich::vcovHC(structural_equation_2, type='HC2'))

stargazer(structural_equation_1_robust, structural_equation_2_robust)
```

The following outputs are the two structural equations that we have computed. We can see that the outcome is sign invariant with respect to the empirical lambda that we use and with the learner used to compute $\hat {log(Wage)}$.

## References

James, G., Witten, D., Hastie, T., & Tibshirani, R. (2013). An introduction to statistical learning. Springer.

Mroz, T. A. (1984). The sensitivity of an empirical model of married women's hours of work to economic and statistical assumptions. Stanford University.

Paliouras, G., Jessen, H. C., & Athens, U. K. (1999). Statistical and learning approaches to nonlinear modeling of labour force participation. Neural Network World, 9, 341-364.

Weerasooriya, T. R. (2018). High-dimensional sample selection models: Machine-learning algorithms in the Heckman two-step.
